Application of the least-squares method to the multiscale transport equation in porous media

### 2011-03-20 MSTE_methodology.tex
con archivo de nomenclaturas

\usepackage[normalem]{ulem}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%This template is created for complying with the author instructions for the 
%8th International Conference on CFD in Oil & Gas, Metallurgical and Process Industries
%hosted by SINTEF/NTNU, Trondheim Norway
%21-23 June 2011
%
%Sverre G. Johnsen (sverre.g.johnsen@sintef.no)
%SINTEF Materials and Chemistry

@@

\newcommand{\TODO}[1]{\textcolor{blue}{TODO: #1} \\}
\newcommand{\Fede}[1]{\textcolor{green}{Fede: #1} \\}
\newcommand{\Pablo}[1]{\textcolor{blue}{#1}}
\newcommand{\newf}[1]{\textcolor{green}{#1}}
\newcommand{\nof}[1]{\textcolor{cyan}{{\sout{#1}}}}

@@

Studies on the water-soil characteristic curve \citep{durner1994}, experimental characterization techniques such as gas adsorption or mercury porosimetry \citep{Navas,Jaroniec1997,Abell1999}\nof{(nitrogen adsorption and mercury)}\newf{,} and the popularization of pore networks \citep{zhang1994, held2001, blunt2002, piri2005, ahrenholz2008, sholokhova2009} have \nof{solidify}\newf{consolidated} the concept of coexisting \nof{different-size }pores\nof{, better represented as the} \newf{of different sizes, which can be better represented with a} pore size distribution, e.g. see \cite{dullien1991}.

@@

The multi-scale transport equation for porous media (MSTE) was proposed as an extension of approaches considering a finite number of groups as studied for example by \cite{Chen1989}, \cite{bouffard2001}. \cite{DupuySchwarz} introduced one internal coordinate to the continuity problem for accounting the multi scale transport of species and the scale-dependent physics. Furthermore, the MSTE defines a inter-pore redistribution function $h$ that makes the equation open and requires further research to find the most adequate closure or model.

@@

The term \nof{pore-size have}\newf{pore size has} been used loosely in the literature and needs to be addressed first. \newf{In our framework, it is a characteristic length that represents the pore scale. }\cite{Dullien} proposed \nof{to use}\newf{to define the pore size as} twice the hydraulic radius\nof{as a convenient definition but others might work as well in the present approach since what we are after is a characteristic length that will represent the pore scale}. \nof{Any of these choices requires a proper definition of a pore as a unity, for which any arbitrary definition of a portion of the void volume can chosen where a pore size can be measure (e.g. hydraulic radius) for each pore.}\newf{Other definitions may be used. Regardless of the choice, a consitent definition of a pore as a unity is required.} Most pore networks agree on the definition of these elementary units. Each pore will therefore have a volume. In the MSTE, all the void volume is assigned to pores, and the concept of throat is not needed\newf{. This is different to} \nof{making a difference with }some pore network strategies.

@@

\TODO{    * Point out challenges for solving the equation}

@@

>>>>

%\usepackage[normalem]{ulem}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%This template is created for complying with the author instructions for the 
%8th International Conference on CFD in Oil & Gas, Metallurgical and Process Industries
%hosted by SINTEF/NTNU, Trondheim Norway
%21-23 June 2011
%
%Sverre G. Johnsen (sverre.g.johnsen@sintef.no)
%SINTEF Materials and Chemistry

@@

\newcommand{\TODO}[1]{\textcolor{blue}{TODO: #1} \\}
\newcommand{\Fede}[1]{\textcolor{green}{Fede: #1} \\}
\newcommand{\Pablo}[1]{\textcolor{blue}{#1}}
\newcommand{\newf}[1]{\textcolor{green}{#1}}
\newcommand{\nof}[1]{\textcolor{cyan}{}}
%\newcommand{\nof}[1]{\textcolor{cyan}{{\sout{#1}}}}

@@

Studies on the water-soil characteristic curve \citep{durner1994}, experimental characterization techniques such as gas adsorption or mercury porosimetry \citep{Navas,Jaroniec1997,Abell1999}\nof{(nitrogen adsorption and mercury)}\newf{,} and the popularization of pore networks \citep{zhang1994, held2001, oren2002, blunt2002, piri2005, ahrenholz2008, sholokhova2009} have \nof{solidify}\newf{consolidated} the concept of coexisting \nof{different-size }pores\nof{, better represented as the} \newf{of different sizes, which can be better represented with a} pore size distribution, e.g. see \cite{dullien1991}.

@@

The multi-scale transport equation for porous media (MSTE) was proposed as an extension of approaches considering a finite number of groups as studied for example by \cite{Chen1989}, \cite{bouffard2001}. One internal coordinate was recently introduced \citep{DupuySchwarz} to the continuity problem for accounting the multi-scale transport of species and the scale-dependent physics. Furthermore, the MSTE defines a inter-pore redistribution function $h$ that makes the equation open and requires further research to find the most adequate closure or model.

@@

The term \nof{pore-size have}\newf{pore size has} been used loosely in the literature and needs to be addressed first. \newf{In our framework, it is a characteristic length that represents the pore scale. }\cite{Dullien} proposed \nof{to use}\newf{to define the pore size as} twice the hydraulic radius\nof{as a convenient definition but others might work as well in the present approach since what we are after is a characteristic length that will represent the pore scale}. \nof{Any of these choices requires a proper definition of a pore as a unity, for which any arbitrary definition of a portion of the void volume can chosen where a pore size can be measure (e.g. hydraulic radius) for each pore.}\newf{Other definitions may be used. Regardless of the choice, a consitent definition of a pore as a unity is required.} Most pore networks agree on the definition of these elementary units. Each pore will therefore have a volume. In the MSTE, all the void volume is assigned to pores, and the concept of throat \citep{oren2002, blunt2002, piri2005} is not needed\newf{. This is different to} \nof{making a difference with }some pore network strategies.

@@

%\TODO{    * Point out challenges for solving the equation}

@@

The combination of the MSTE with LSQ enables a set of new physics that now can be modelled. We have shown how the specific discharge, and therefore, the advecting velocity can be different at different scales, having travelling times between the smallest and the largest pore scales that vary two order of mangitude. The cumulative distribution function of any variable, such as the tracer concentration can be recovered at any point of the domain. Furthermore, having a source term that works as sink and as source depending on the pore scale is also possible enabling the modelling of the competition between heterogeneous rections and reactions occuring in the liquid bulk. Finally, the incorporation of a redistribution tems add the complexity of an integral term that binds physics happening at different scale loosing the locality behavior proper of a normal transport equation and still this can be fully captured by the LSQ also with exponential convergency.
